---
title: "RProgramming_Week1"
author: "Luis Ivan"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1 - R objects and attributes

5 basic types of objects:  
-  Character  
-  Numeric  
-  Integer  
-  Complex  
-  Boolean  

Special numbers: **Inf** (Infinite) , **NaN** (Not a Number)

Objects have attributes

# Week 1 - Vectors and lists

Different ways of creating vectors of objects:

```{r cars}
x <- c(0.5, 0.6)
x <- c(TRUE, FALSE)
x <- c(T, F)
x <- c("a", "b", "c")
x <- 9:29
x <- c(1+0i, 2+4i)

x <- vector("numeric", length = 10)
```

**Coercion** is similar to casting. Vectors of integer and character will coerce to character.

**Explicit coercion** happens with as.* functions.
```{r coercion}
x<- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

##Lists

*Lists* are vectors with elements of different classes. ~~Similar to structs in C~~.
They have double brackets, so they are identifiable that way.
Vectors have single bracket.

```{r lists}
x<-list(1, "a", TRUE, 1+4i)
x
```

# Matrices

```{r matrices}
m <- matrix(nrow = 2, ncol = 3)
m
attributes(m)
```

Constructed column-wise. Upper left running down the columns.
Can be constructed also from a vector.

```{r from vector}
m <- 1:10
m
dim(m) <- c(2,5)
m
```

Can be created by binding two variables with cbind(var1, var2) or rbind(x, y). *Rowbind or columnbind*

# Factors

Special type of vector for categorical data.  
-  Integer vector where each integer has a label.  
-  Treated by modelling functions like lm() and glm().  
-  Something like enums in C.

```{r factors}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x)
unclass(x)
```

Order of levels can be set by using argument factor(). Default is alphabetical order

```{r order factors}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x
```