---
title: "RProgramming_Week2"
author: "Luis Ivan"
date: "11/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Control Structures in R

- if, else  
- for  
- while: Execute a loop **while** a condition is true  
- repeat: execute an infinite loop  
- break: break execution of loop  
- next: skip iteration of loop

### if - else if - else

You can write the if directly in the declaration, or normally like in C.
y <- if(x > 3) {
	10
 }
else
{
0
}

### for loop

for(i in 1:10)
{
<condition>
}

for(i in seq_along(x)){ } //loops for length of vector but i is integer

for(letter in x){ } //loops for length of vector but letter is char

### while loop

Same as in C, can be stuck in infinite loops since there's no hardcoded exit condition

### Repeat / break

repeat initiates an infinite loop, stops with a *break*. Not recommendable since it's an
infinite loop just like the while.

### next

Skips an iteration on a for loop / while loop / ...

# Functions

User of R -> Programmer of R
Functions are R objects of Class function. Can be passed as arguments, and can be nested.
Named arguments where potentially they have default values.
Arguments can be called *positionally* or *by name*.

**If arguments are matched by name, they are taken out of the list. The unnamed ones
are matched in the order they are listed (excluding the previously named)**

Arguments are only evaluated as needed. -> Lazy evaluation
```{r lazy evaluation}
f <- function(a,b){
  print(a)
}
f(54)
```

'...' argument is used to pass many undefined arguments (not want to copy full
set of arguments). Also necessary when number of arguments passed to function
cannot be known in advance. Cannot be matched positionally neither anything after.

# Binding

When R binds a value to a symbol, it searches first in the global environment for a
symbol matching the requested one. If not, it looks in packages in search list. 
The order is in the following command:

```{r search}
search()
```

When loading a new library, it gets put in position 2 in the search list by default.
R has separate namespaces for functions and non-functions. (can create a var and a func with same name).

### Scoping rules

Determine how a value is associated with a free variable in a function. Uses static scoping, instead of dynamic scoping.
Lexical scoping (or static) eases statistical operations.
Lexical scoping means that free variables are searched in the environment in which the function was defined.  
-  If the variables are not in the environment, they are looked for in the parent environment.  
- If it's not in the top-level environment, it looks in the next element in the search list, until we hit the **empty environment**. Then, error.

We can have functions in another function!! 

```{r lexical scoping}
make.power <- function(n){
  pow <- function(x){
    x^n
  }
  pow
}

cube <- make.power(3) #Cube is pow with x^3
square <- make.power(2) #Is pow with x^2
cube(3)
square(3)
```

Seems kind of like a virtual function in C++.

Due to this scoping, all objects are stored in memory. All functions carry a pointer to the defining environment.

### Benefit of objective functions

No need to carry argument lists, since the obj functions can be built which contain all the necessary data for evaluating the function.

# Coding Standards

The standards are:  
- Indent your code  
- Limit the width of the code (80 columns circa)  
- **Limit the length of the functions (one activity)**

# Dates and Times

Special topic. Date represented with Date class. Time represented by POSIXct class.
Dates stored as days from 1970-01-01.
Times stored as seconds since 1970-01-01.

```{r dates}
x <- as.Date("1970-01-01")
x
unclass(x)
```

### Times in R
- POSIXct : just a large integer  
- POSIXlt : Stores info as day of week, month, day of month  

Useful funcs as weekdays, months, quarters.
```{r times}
x <- Sys.time()
typeof(x)
x
unclass(x)
p<-as.POSIXlt(x)
names(unclass(p))
p$yday
```
