---
title: "RProgramming_Week2"
author: "Luis Ivan"
date: "11/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Control Structures in R

- if, else  
- for  
- while: Execute a loop **while** a condition is true  
- repeat: execute an infinite loop  
- break: break execution of loop  
- next: skip iteration of loop

### if - else if - else

You can write the if directly in the declaration, or normally like in C.
y <- if(x > 3) {
	10
 }
else
{
0
}

### for loop

for(i in 1:10)
{
<condition>
}

for(i in seq_along(x)){ } //loops for length of vector but i is integer

for(letter in x){ } //loops for length of vector but letter is char

### while loop

Same as in C, can be stuck in infinite loops since there's no hardcoded exit condition

### Repeat / break

repeat initiates an infinite loop, stops with a *break*. Not recommendable since it's an
infinite loop just like the while.

### next

Skips an iteration on a for loop / while loop / ...

# Functions

User of R -> Programmer of R
Functions are R objects of Class function. Can be passed as arguments, and can be nested.
Named arguments where potentially they have default values.
Arguments can be called *positionally* or *by name*.

**If arguments are matched by name, they are taken out of the list. The unnamed ones
are matched in the order they are listed (excluding the previously named)**

Arguments are only evaluated as needed. -> Lazy evaluation
```{r lazy evaluation}
f <- function(a,b){
  print(a)
  print(b)
}
f(54)
```

'...' argument is used to pass many undefined arguments (not want to copy full
set of arguments). Also necessary when number of arguments passed to function
cannot be known in advance. Cannot be matched positionally neither anything after.