---
title: "RProgramming_Week1"
author: "Luis Ivan"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1 - R objects and attributes - DataTypes

5 basic types of objects:  
-  Character  
-  Numeric  
-  Integer  
-  Complex  
-  Boolean  

Special numbers: **Inf** (Infinite) , **NaN** (Not a Number)

Objects have attributes

## Vectors and lists

Different ways of creating vectors of objects:

```{r cars}
x <- c(0.5, 0.6)
x <- c(TRUE, FALSE)
x <- c(T, F)
x <- c("a", "b", "c")
x <- 9:29
x <- c(1+0i, 2+4i)

x <- vector("numeric", length = 10)
```

**Coercion** is similar to casting. Vectors of integer and character will coerce to character.

**Explicit coercion** happens with as.* functions.
```{r coercion}
x<- 0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

##Lists

*Lists* are vectors with elements of different classes. ~~Similar to structs in C~~.
They have double brackets, so they are identifiable that way.
Vectors have single bracket.

```{r lists}
x<-list(1, "a", TRUE, 1+4i)
x
```

## Matrices

```{r matrices}
m <- matrix(nrow = 2, ncol = 3)
m
attributes(m)
```

Constructed column-wise. Upper left running down the columns.
Can be constructed also from a vector.

```{r from vector}
m <- 1:10
m
dim(m) <- c(2,5)
m
```

Can be created by binding two variables with cbind(var1, var2) or rbind(x, y). *Rowbind or columnbind*

## Factors

Special type of vector for categorical data.  
-  Integer vector where each integer has a label.  
-  Treated by modelling functions like lm() and glm().  
-  Something like enums in C.

```{r factors}
x <- factor(c("yes", "yes", "no", "yes", "no"))
x
table(x)
unclass(x)
```

Order of levels can be set by using argument factor(). Default is alphabetical order

```{r order factors}
x <- factor(c("yes", "yes", "no", "yes", "no"),
            levels = c("yes", "no"))
x
```

## Missing values

Verified by is.na() or is.nan()
NaN value is also NA, but not the other way around.
NA is a numeric missing value, NaN is not a number.

## Data Frames

Tabular data, tables. Every element = same length. Column can be different types.

Read by read.table() or read.csv().
data.matrix to convert to matrix

```{r data frames}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F), temp = c("Lu", "Da", "Mi", "No"))
x
nrow(x)
ncol(x)
```

## Names attributes

Can add names to describe objects to lists, frames, matrices

```{r names}
x <- data.frame(foo = 1:4, bar = c(T, T, F, F), temp = c("Lu", "Da", "Mi", "No"))
y <- list(a = 1, b = 2, c = 3)
names(x) <- c("foo", "bar", "nor")
x
y
```

# Reading Data

read.table Arguments:  
-  File: name of file  
-  Header: if the file has header file  
-  Sep: how columns are separated. Default is *space*.  
-  colClasses: class of each column  
-  nrows: number of rows  
-  comment.char: how it's a comment  
-  skip: number of lines from beginning  
-  stringsAsFactors: should char coded as factors?  

*Read the help page for read.table()*
Memory (MB) is Rows*Columns*SizeOfData / (2to20).

# Textual formats

dumping and dputing - textual format is editable.
Read about dump and dput/dget. The important stuff is that it saves an .R file with the value and metadata of programs.

# Interfaces to outside world

file - normal file **(r - readonly / w - writing)**
gzfile - gzip file
bzfile - bzip2
url - webpage

# Subsetting R Objects

[ returns an object of the same class
[[ extracts elements of lists or data frames.
$ obtains elements by name.

```{r na values}
x <- c(1, 2, NA, 4, NA, 6)
bad <- is.na()
x[!bad]

x <- c(1, 2, NA, 4, NA, 6)
y <- c("a",  "b", NA, "d", NA, NA)

good <- complete.cases(x,y)
good
x[good]
y[good]
```