---
title: "RProgramming_Week3"
author: "Luis Ivan"
date: "11/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions

## lapply

Loop over a list and evaluate a function for each element.
Two arguments:  
- a list (can be coerced as as.list)  
- a function  

```{r lapply}
x <- list(a=1:5, b=rnorm(10))
lapply(x, mean)

y <- list(a=1:4, b=rnorm(10), c=rnorm(20,1), d=rnorm(100,5))
lapply(y, mean)

z<-1:4
lapply(z, runif, min=0, max=10)
```

Now with matrices
```{r lapplyMatrix}
x<-list(a = matrix(1:4, 2, 2), b=matrix(1:6, 3, 2))
x
lapply(x, function(elt) elt[,1])
```

## sapply
Try to simplify lapply if possible.
If every element in the result list is length 1, returns a vector.
If the result is a list where all elements are vectors of same length, returns a matrix.
If it's mixed, then a list is returned.

```{r sapply}
x <- list(a=1:4, b=rnorm(10), c=rnorm(20,1), d=rnorm(100,5))
lapply(x, mean)
sapply(x, mean)
```

## apply

Used to evaluate function over margins of array.
Used to apply a function to rows or columns of matrix
It's not faster than writing a loop, but it's a one liner!

Three arguments:  
-  Array
- Margin (which margin is retained)  
- Function to be applied

```{r apply}
x <- matrix(rnorm(200), 20, 10)
apply(x,2,mean) #columnwise
apply(x,1,mean) #rowwise
```

Functions for simple functions such as sum and mean, we have **rowsums**, **rowmeans**, **colsums**, **colmeans**.

```{r quantiles}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))
```

## mapply

Multivariate apply, which applies a function in parallel over set of arguments.

Example.
Instead of typing
*list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))*
we can type
``` {r mapply}
mapply(rep, 1:4, 4:1)

noise <- function(n, mean, sd){
  rnorm(n,mean,sd)
}
noise(5,1,2)
noise(1:5, 1:5, 2) #incorrect
mapply(noise, 1:5, 1:5, 2) #correct
```

## tapply

apply a function over subsets of vectors.
Arguments:
- a Vector
- an index is a factor or list of factors
- Function to be applied
- Other arguments of function
- simplify = should the result be simplified (**sapply**)

```{r tapply}
x <- c(rnorm(10), runif(10), rnorm(10,1))
f <- gl(3, 10)
tapply(x, f, mean)
tapply(x, f, mean, simplify = FALSE)
```

## split

Takes a vector and splits it into groups determined by factors
Arguments:  
-  Vector, list or dataframe  
-  factor or list of factors  
-  drop = empty factors should be dropped?  
  
```{r split}
x <- c(rnorm(10), runif(10), rnorm(10,1))
f <- gl(3, 10)
split(x, f)
lapply(split(x,f), mean)
```

Previous example can be done also with tapply.

```{r complicatedSplit}
library(datasets)
head(airquality)
s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[,c("Ozone", "Solar.R", "Wind")]))
sapply(s, function(x) colMeans(x[,c("Ozone", "Solar.R", "Wind")]))
sapply(s, function(x) colMeans(x[,c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
```

You can split in more than one level.

```{r multilevelSplit}
x <- rnorm(10)
f1 <- gl(2,5)
f2 <- gl(5,2)
interaction(f1, f2)
str(split(x, list(f1,f2)))
str(split(x, list(f1,f2), drop=TRUE))
```